---
title: "Potential_WTA"
author: "Michael Mann"
date: "12/22/2021"
output: html_document
---
# peak rank at 100 win rate and career slam semifinal appearences.


# factors:
# rank at age 20 * win percentage at age 20
# parse out by surface



# to be fair since some players are younger and haven't had a chance to win a slam. So going to cut it off after 30. 
# I will limit my search between players who were 20 in 1970 to 2010. 

# thus have to be born between 1950 and 1990. 


### Tasks

# merge all match data into one dataframe. Clean it up. 
# Calculate win percentage for players between 1970 and 2010. Then filter it for just players who were 20. 
# tally up number of slam wins and semi-final appearences. 

# merge two sources of data into one dataframe. 

```{r packages}
library(tidyverse)

```
Creates list of all players and their matches between 1980 and 2019
```{r import and merge all data}
wta_data <- 
  tibble(
    files = 
      c(list.files(path = "tennis_wta-master", pattern = "wta_matches_198..csv", full.names = T),
      list.files(path = "tennis_wta-master", pattern = "wta_matches_199..csv", full.names = T),
      list.files(path = "tennis_wta-master", pattern = "wta_matches_200..csv", full.names = T),
      list.files(path = "tennis_wta-master", pattern = "wta_matches_201..csv", full.names = T), 
       list.files(path = "tennis_wta-master", pattern = "wta_matches_2021.csv", full.names = T))
    ) %>% 
  mutate(matches = map(files, read_csv, 
                       col_types =  
                         cols_only(
                            tourney_id = col_character(),
                            tourney_name = col_character(),
                            surface = col_character(),
                            tourney_level = col_character(),
                            tourney_date = col_double(),
                            match_num = col_double(),
                            winner_id = col_double(),
                            winner_entry = col_character(),
                            winner_name = col_character(),
                            winner_hand = col_character(),
                            winner_ht = col_double(),
                            winner_ioc = col_character(),
                            winner_age = col_double(),
                            loser_id = col_double(),
                            loser_entry = col_character(),
                            loser_name = col_character(),
                            loser_hand = col_character(),
                            loser_ht = col_double(),
                            loser_ioc = col_character(),
                            loser_age = col_double(),
                            score = col_character(),
                            best_of = col_double(),
                            round = col_character()   
                          )
                       )
         ) %>% # this import specfic columns for each datasheet
  mutate(year = str_extract(string = files, pattern = "[:digit:][:digit:][:digit:][:digit:]")) %>%
  mutate(year = as.numeric(year)) %>%
  unnest(matches)
  

```



creates player_record_year which has the number of wins per year for each player. 
```{r player record}
# winner table
wins_player_year <- 
  wta_data %>%
  group_by(winner_name, winner_id,  year) %>%
  summarise(matches_won = n(), .groups = "drop") %>%
  drop_na() %>%
  rename(name = winner_name)


losses_player_year <- 
  wta_data %>%
  group_by(loser_name, loser_id, year) %>%
  summarise(matches_loss = n(), .groups = "drop") %>%
  drop_na() %>%
  rename(name = loser_name)

player_record_year <- 
  wins_player_year %>%
  full_join(losses_player_year, by = c("name", "year")) %>%
  mutate(total_played = matches_won + matches_loss) %>%
  mutate(percent_won = 100 * (matches_won /total_played ))
```



Creates a list of all players who were 20 between 1980 to 2010. 
```{r potential players list}
players <- read_csv(file = "tennis_wta-master/wta_players.csv")

potential_players <- 
  players %>%
  mutate(dob = lubridate::ymd(dob)) %>%
  drop_na(dob) %>%
  mutate(year_born = lubridate::year(dob)) %>%
  mutate(year_at_20 = year_born + 20) %>% # calculate the year they turned 20
  filter(year_at_20 >= 1980 & year_at_20 <= 2010) 
  
  

```


```{r players who entered top 100 and within the time window}

ranking_data_year_low <- 
  tibble(
    files = 
      c(list.files(path = "tennis_wta-master", pattern = "wta_rankings_.0", full.names = T))) %>% 
  mutate(rankings = map(files, read_csv)) %>%
  unnest(rankings) %>%
  mutate(ranking_date= lubridate::ymd(ranking_date)) %>%
  mutate(year = lubridate::year(ranking_date)) %>%
  group_by(year, player) %>%
  summarise(top_rank_by_year = min(rank), .groups = "drop") %>%
  rename(player_id = player) %>%
  right_join(potential_players, by =  "player_id") %>%
  filter(year_at_20 == year) %>%
  filter(top_rank_by_year <= 100)

top_100_women <- 
  ranking_data_year_low %>%
  pull(player_id)
```



# makes a dataframe with the number of slam appearences and finals by player.
```{r compile records of players}
slam_sucess <- 
  wta_data %>%
  select(winner_id, loser_id, tourney_name, round) %>%
  filter(tourney_name %in% c("Australian Open", "Roland Garros", "Wimbledon", "US Open")) %>%
  filter(round == "F"| round == "SF") %>%
  pivot_longer(cols = c("winner_id", "loser_id"), names_to = "Outcome", values_to = "player_id") %>%
  mutate(Outcome = str_replace(Outcome, pattern = "_id", replacement = "")) %>%
  group_by(round, player_id, Outcome) %>%
  count() %>%
  pivot_wider(names_from = "round", values_from = "n") %>%
  ungroup() %>%
  replace_na(list("F"=  0,"SF" = 0)) %>%
  group_by(player_id) %>%
  mutate(Finals_Appearences = sum(`F`), Semifinals_Apperences = sum(SF)) %>%
  ungroup() %>%
  filter(Outcome == "winner") %>%
  rename(Finals_Won = `F`, Semifinals_Won = "SF") %>%
  select(-Outcome)



```

# this puts all the pieces together. 
```{r }
processed_dataset <- 
  player_record_year %>%
  filter(winner_id %in% top_100_women) %>%
  rename(player_id = winner_id) %>%
  left_join(potential_players, by = "player_id") %>%
  filter(year_at_20 == year) %>% # this is a different filter than above since this is the year of their winning record. 
  left_join(slam_sucess, by = "player_id") %>%
  select(name, year, year_born, year_at_20, matches_won, matches_loss, total_played, percent_won, Finals_Won:Semifinals_Apperences) %>%
  replace_na(list("Finals_Won"=  0,"Semifinals_Won" = 0, "Finals_Appearences"  = 0, "Semifinals_Apperences" = 0))
  
  


```


```{r preliminary plots}

processed_dataset %>%
  filter(total_played > 20) %>%
  ggplot(aes(x = percent_won, y = Finals_Appearences, label = name)) +
  geom_text()


```


```{r }
library(emmeans)
processed_dataset_20 <- 
 processed_dataset %>%
  filter(total_played > 20) %>%
  mutate(slam_champ = Finals_Won > 0) %>%
  mutate(slam_semifinalist = Semifinals_Apperences > 0)

m1 <- glm(slam_champ ~ percent_won, family = binomial(), data = processed_dataset_20)
res <- DHARMa::simulateResiduals(m1, plot = T)

null <- glm(Finals_Appearences ~ 1, family = binomial(), data = processed_dataset_20)
anova(m1, null, test="Chisq")
m1_grid <- emmeans::ref_grid(m1, at =list(percent_won = seq(from = 50, to = 90, by = 2.5)))
model_predict <- plot(m1_grid)
model_predict + scale_x_continuous(breaks = c(-8:4)) + geom_vline(xintercept = 0) 


```
```{r }
library(ggridges)


processed_dataset_20 %>%
  ggplot(aes(y = slam_champ, x = percent_won)) +
  geom_density_ridges(alpha = .2,) + geom_vline(xintercept = 65.6) + 
  theme_ridges(grid = FALSE, center_axis_labels = TRUE)

```



```{r upcoming playres}

upcoming_players <- 
  players %>%
  mutate(dob = lubridate::ymd(dob)) %>%
  drop_na(dob) %>%
  mutate(year_born = lubridate::year(dob)) %>%
  mutate(year_at_20 = year_born + 20) %>% # calculate the year they turned 20
  filter(year_at_20 >= 2018) 

player_record_year %>%
  rename(player_id = winner_id) %>%
  left_join(upcoming_players, by = "player_id") %>%
  drop_na(year_born) %>%
  filter(year_at_20 == year -1) %>%
  filter(total_played >= 20) %>%
  View()


```